Well-quasi-orderings on word languages

### 2024-10-18 src/amalgam.tex
adding more and more todos

\begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
    \proofref{infix-amalgamation:thm}
    Let $L \subseteq \Sigma^*$ be a language recognized by an 
    \kl{amalgamation system}.
    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
    \kl(language){bounded}.
\end{theorem}

\begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
% 	\label{infix-wqo-is-emptiness:thm}
	Let $\mathcal{C}$ be a class of languages recognized by 
    \kl{amalgamation systems}
    and closed under
    \kl{rational transductions}. 
    Then the following are equivalent:
	\begin{enumerate}
        \item\label{wqo-infix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{infix relation} is decidable for languages in $\mathcal{C}$.
        \item\label{wqo-prefix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{prefix relation} is decidable for languages in $\mathcal{C}$.
        \item\label{emptiness-decidable} Emptiness is decidable for languages in $\mathcal{C}$.
	\end{enumerate}
\end{theorem}

\subsection{Amalgamation Systems}
\label{infixes-amalgamation:subsec}

@@

For the next result, we impose a mild restriction by requiring the language
class to be closed under \kl{rational transductions}. This property holds for
most common language classes, including all known \kl{amalgamation systems}.

\AP A class of languages $\mathcal{C}$ is \intro{recognized by amalgamation
systems} when, for every language $L \in \mathcal{C}$, there exists an
\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ such that $L$ is the
language of this system.

\AP A \intro{rational transduction} is a function $f \colon \Sigma^* \to
\Gamma^*$ that can be defined by a finite automaton with output, or
equivalently by a finite set of rules of the form $u \to v$ where $u \in
\Sigma^*$ and $v \in \Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the
image of a language $L$ by a rational transduction $f$. A class of languages
$\mathcal{C}$ is \intro{closed under rational transductions} when, for every
language $L \in \mathcal{C}$, and every rational transduction $f \colon
\Sigma^* \to \Gamma^*$, the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ...}

@@

    \cref{wqo-infix-decidable} $\Rightarrow$ \cref{wqo-prefix-decidable}. We
    just consider the transduction

>>>>

\begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
    \proofref{infix-amalgamation:thm}
    Let $L \subseteq \Sigma^*$ be a language recognized by an 
    \kl{amalgamation system}.
    If $L$ is well-quasi-ordered by the infix relation then $L$ is
    \kl(language){bounded}.
\end{theorem}

\todo[inline]{There are missing effective restrictions in the statement of the theorem: we should
have descriptions of the languages, (effective) stability under transductions etc.}

\begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
% 	\label{infix-wqo-is-emptiness:thm}
	Let $\mathcal{C}$ be an \kl{effective amalgamative class} of languages
    closed under
    \kl{rational transductions}. 
    Then the following are equivalent:
	\begin{enumerate}
        \item\label{wqo-infix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{infix relation} is decidable for languages in $\mathcal{C}$.
        \item\label{wqo-prefix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{prefix relation} is decidable for languages in $\mathcal{C}$.
        \item\label{emptiness-decidable} Emptiness is decidable for languages in $\mathcal{C}$.
	\end{enumerate}
\end{theorem}

\subsection{Amalgamation Systems}
\label{amalgamation-systems:subsec}

@@

\AP For the next result, we impose a mild restriction by requiring the language
class to be closed under \kl{rational transductions}. This property holds for
most common language classes, including all known \kl{amalgamation systems}. A
\intro{rational transduction} is a function $f \colon \Sigma^* \to \Gamma^*$
that can be defined by a finite automaton with output, or equivalently by a
finite set of rules of the form $u \to v$ where $u \in \Sigma^*$ and $v \in
\Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the image of a language
$L$ by a rational transduction $f$. A class of languages $\mathcal{C}$ is
\intro{closed under rational transductions} when, for every language $L \in
\mathcal{C}$, and every rational transduction $f \colon \Sigma^* \to \Gamma^*$,
the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ... Do we allow
the alphabet to vary?}

\AP A class $\mathcal{C}$ of languages is an \intro{effective amalgamative
class} whenever there exists a finite presentation of the languages \dots\todo{
define formally}. And for every $L \in \mathcal{C}$, there exists an
\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ such that $L$ is the
language of this system.

\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
Note that this implies boundedness and simultaneous boundedness of languages in
$\mathcal{C}$ is decidable \cite{ASZZ24}. \todo{define formally}

@@

    \cref{wqo-infix-decidable} $\Rightarrow$ \cref{wqo-prefix-decidable}. We
    just consider the transduction \todo{fresh symbols? can we assume we have those??}

### 2024-10-18 src/conclusion.tex
adding more and more todos

>>>>

\subparagraph{Lexicographic orderings} There is another natural ordering on
words, the \emph{lexicographic ordering}, which does not fit well in our
current framework because it is always of \kl{ordinal width} $0$. However, the
order-type of the lexicographic ordering over \kl{regular languages} has
already been investigated in the context of infinite words, and it would be
interesting to see if one can extend these results to decide whether such an
ordering is \kl{well-founded} for languages recognized by \kl{amalgamation
systems}. \todo{Find references for that?}

### 2024-10-18 src/infixes.tex
adding more and more todos

\AP In this section, we study languages equipped with the \kl{infix relation}.
As opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix
relation} can lead to very complicated \kl{well-quasi-ordered} languages.
Formally, the upcoming \cref{infix-embedding:thm} shows that
\emph{any} countable partial-ordering with finite initial segments can be
embedded into the infix relation of a language. To make the former statement
precise, let us recall that an \intro{order embedding} from a quasi-ordered set
$(X, \preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon
X \to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
\preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq
y$ and $y \preceq x$ implies $x = y$. 
\begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
    \label{infix-embedding:thm}
    Let $(X, \preceq)$ be a \kl{partially ordered} set,
    and $\Sigma$ be an alphabet with at least two letters.
    Then the following are equivalent:
    \begin{enumerate}
        \item \label{infix-embedding-embeds:item} 
            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
        \item \label{infix-embedding-count:item}
            $X$ is countable, and for every $x \in X$,
            its \kl{downwards closure}
            $\dwset[\preceq]{x}$ is finite.
    \end{enumerate}
\end{lemma}
\begin{figure}
    \centering
    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
        inside the \kl{infix relation} for $\set{a,b,\#}^*$
        using the encoding of \cref{infix-embedding:thm}, restricted to words
        of length at most $3$. To obtain smaller words,
        we replaced $D_i$ by $\max D_i$ in this construction.
    }
    \label{infix-embedding:fig}
\end{figure}

@@

\begin{corollary}
    \label{inf-period-union-chains:lem}
    Let $x,u,y \in \Sigma^*$ be words.  The following 
    are finite unions of \kl{grids} for the infix relation:
    $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
    and $\InfPeriodChain{x} u \InfPeriodChain{y}$.
\end{corollary}
\begin{proof}
    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
    relation (rsing \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
    for the \kl{suffix relation},
    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
    (using again \cref{inf-period-chain:lem}).
    As a consequence,
    $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
    of \kl{grids} for the \kl{infix} relation.
\end{proof}

@@

    Assume by contradiction that $L$ is \kl{bounded}. In this case, there exist
    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
    because $u \infleq \ThueMorse$, which is \kl{cube-free}.

@@

    Now, let us prove that these bounds are tight. It is clear that
    $\oHeight{L} = \omega$: given any number $n \in \Nat$, one can construct a
    \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
    considering the first $n$ prefixes of the \kl{Thue-Morse sequence} by
    decreasing size.
    Let us now prove that $\oWidth{L} = \omega$. Assume by contradiction that
    $\oWidth{L}$ is finite. Then, $L$ can be written as a finite union of
    \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
    \kl{bounded}, which is absurd by \cref{thue-morse:lemma}.
    Finally, because the \kl{ordinal width} is at most the \kl{maximal order
    type}, we conclude that the \kl{maximal order type} of $L$ is also $\omega$.
\end{proof}

>>>>

\AP In this section, we study languages equipped with the \kl{infix relation}.
As opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix
relation} can lead to very complicated \kl{well-quasi-ordered} languages.
Formally, the upcoming \cref{infix-embedding:thm} shows that
\emph{any} countable partial-ordering with finite initial segments can be
embedded into the infix relation of a language. To make the former statement
precise, let us recall that an \intro{order embedding} from a quasi-ordered set
$(X, \preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon
X \to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
\preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq
y$ and $y \preceq x$ implies $x = y$. 
The embedding defined in \cref{infix-embedding:thm} is illustrated
for the \kl{subword relation} in \cref{infix-embedding:fig}.
\begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
    \label{infix-embedding:thm}
    Let $(X, \preceq)$ be a \kl{partially ordered} set,
    and $\Sigma$ be an alphabet with at least two letters.
    Then the following are equivalent:
    \begin{enumerate}
        \item 
            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
        \item 
            $X$ is countable, and for every $x \in X$,
            its \kl{downwards closure}
            $\dwset[\preceq]{x}$ is finite.
    \end{enumerate}
\end{lemma}
\begin{figure}
    \centering
    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
        inside the \kl{infix relation} for $\set{a,b,\#}^*$
        using the encoding of \cref{infix-embedding:thm}, restricted to words
        of length at most $3$. To obtain smaller words,
        we replaced $D_i$ by $\max D_i$ in this construction.
    }
    \label{infix-embedding:fig}
\end{figure}

@@

@@

    Assume by contradiction that $L$ is \kl(language){bounded}. In this case, there exist
    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
    because $u \infleq \ThueMorse$, which is \kl{cube-free}.

@@

    Now, let us prove that these bounds are tight. It is clear that
    $\oHeight{L} = \omega$: given any number $n \in \Nat$, one can construct a
    \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
    considering the first $n$ prefixes of the \kl{Thue-Morse sequence} by
    decreasing size.
    Let us now prove that $\oWidth{L} = \omega$. Assume by contradiction that
    $\oWidth{L}$ is finite. Then, $L$ can be written as a finite union of
    \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
    \kl(language){bounded}, which is absurd by \cref{thue-morse:lemma}.
    Finally, because the \kl{ordinal width} is at most the \kl{maximal order
    type}, we conclude that the \kl{maximal order type} of $L$ is also $\omega$.
\end{proof}

### 2024-10-18 src/introduction.tex
adding more and more todos

\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
runtime of the algorithm is deeply related to the ``complexity'' of the
underlying quasi-order \todo{CITE???}. One way to measure this complexity is to
consider its so-called \kl{ordinal invariants}: for instance, the
\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and
Parikh \cite{dejongh77}, is the order type of the maximal linearization of a
well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
the size of the set. Better runtime bounds were obtained by considering two
other parameters\todo{cite sylvain width/ HDR?}: the \kl{ordinal height}
introduced by Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž
and Thomas \cite{kriz90b}. Therefore, when characterizing
\kl{well-quasi-ordered} languages, we will also be interested in computing
bounds on their \kl{ordinal invariants}. We refer to the
\cref{ordinal-invariants:subsec} for a more detailed introduction to these
parameters and ordinal computations in general. For the moment, let us just
remark that all languages $L \subseteq \Sigma^*$ have an \kl{ordinal height} of
at most $\omega$.

@@

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
	\setcounter{enumi}{3}
    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ 
        is a \kl{chain} for the \kl{suffix relation},
        and where
        each $P_i$ is a \kl{chain} for the \kl{prefix relation}.
\end{enumerate}
}

This result directly translates into bounds on the possible \kl{ordinal invariants}
of such languages similarly as for the \kl{prefix relation}.
{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{4}
    \item All \kl{bounded languages}
            $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
            have an \kl{ordinal height} of at most $\omega$,
            an \kl{ordinal width} strictly below $\omega^2$, and
            a \kl{maximal order type} strictly below $\omega^3$.
\end{enumerate}
}

\AP
Finally, we provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered}, which takes as inputs languages recognized
by \kl{amalgamation systems} \cite{ASZZ24}. Such systems will be
formally introduced in \cref{amalgamation-systems:subsec}, but for the moment
let us just say that they include many classical computational models such as
finite automata, context-free grammars, and petri nets \cite{ASZZ24}. 
{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{5}
    \item There is a decision procedure for checking whether a language $L$ 
        recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
        by the \kl{prefix relation}, the \kl{suffix relation},
        or the \kl{infix relation}. Furthermore, if it is \kl{well-quasi-ordered}, 
        then $L$ is a \kl{bounded language}.
        \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
\end{enumerate}
}
Notice that the above result implies that the hypothesis of
\kl{bounded languages} on the theoretical side is not a restriction at all in
practice.

@@

\subparagraph{Outline} We introduce in \cref{prelims:sec} the
necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
In
\cref{prefixes:sec}, which is relatively
self-contained, we study the \kl{prefix relation} and prove in
\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
languages by the \kl{prefix relation}. In
\cref{infixes-regular:sec}, we
obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
specifically for regular languages
(\cref{infix-finite-automata:thm}). Finally, we generalize this result to all
\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
in
(\cref{infix-amalgamation:thm}).

>>>>

\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
runtime of the algorithm is deeply related to the ``complexity'' of the
underlying quasi-order~\cite{SCHMITZ17}. One way to measure this complexity is
to consider its so-called \kl{ordinal invariants}: for instance, the
\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and Parikh
\cite{dejongh77}, is the order type of the maximal linearization of a
well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
the size of the set. Better runtime bounds were obtained by considering two
other parameters~\cite{SCHMITZ19}: the \kl{ordinal height} introduced by
Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž and Thomas
\cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
languages, we will also be interested in computing bounds on their \kl{ordinal
invariants}. We refer to \cref{ordinal-invariants:subsec}
for a more detailed introduction to these
parameters and ordinal computations in general.

@@

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
	\setcounter{enumi}{3}
    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is (included in) a finite union of languages $S_i \cdot P_i$, where each $S_i$ 
        is a \kl{chain} for the \kl{suffix relation},
        and where
        each $P_i$ is a \kl{chain} for the \kl{prefix relation}.

        \hfill
        (\cref{bounded-language:thm} p.\pageref{bounded-language:thm})
\end{enumerate}
}

This result directly translates into upper bounds on the possible \kl{ordinal invariants}
of such languages similarly as for the \kl{prefix relation}. Notice that these
upper bounds are significantly smaller than \kl{ordinal invariants} of the \kl{subword
relation}.
{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{4}
    \item All \kl{bounded languages}
            $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
            have an \kl{ordinal height} of at most $\omega$,
            an \kl{ordinal width} strictly below $\omega^2$, and
            a \kl{maximal order type} strictly below $\omega^3$.
\end{enumerate}
}

\AP Finally, we provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered}, which takes as inputs languages recognized by
\kl{amalgamation systems} \cite{ASZZ24}. Such systems will be formally
introduced in
\cref{amalgamation-systems:subsec}, but
for the moment let us just say that they include many classical computational
models such as finite automata, context-free grammars, and Petri nets
\cite{ASZZ24}. This provides us with a \emph{meta-algorithm} for deciding
whether a given language is \kl{well-quasi-ordered} by the \kl{prefix
relation}, the \kl{suffix relation}, or the \kl{infix relation} under mild
effective restrictions on the computational model that we call an \kl{effective
amalgamative class}, the formal statement of which we defer to
\cref{infixes-amalgamation-effective:subsec}.

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{5}
    \item 
        Let $L$ be a language recognized by an \kl{amalgamation system}.
        If $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
        then $L$ is a \kl{bounded language}.

        \hfill (\cref{infix-amalgamation:thm} p.\pageref{infix-amalgamation:thm})
    \item 
        Let $\mathcal{C}$ be a \kl{strongly effective amalgamative class} of languages.
        There is a decision procedure that takes as input a language $L \in \mathcal{C}$,
        and decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation},
        the \kl{suffix relation}, or the \kl{infix relation}. 
        \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
\end{enumerate}
}
Let us point out that the above result implies that the hypothesis of
\kl{bounded languages} on the theoretical side is not a restriction at all in
practice. As a practical and more easily understandable 
consequence,
our generic decision procedure applies 
to the class $\mathcal{C}_\text{aut}$ of all languages recognized by finite automata,
and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
which are both \kl{effective amalgamative classes}.

@@

\subparagraph{Outline} 
We introduce in \cref{prelims:sec} the
necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
In
\cref{prefixes:sec}, which is relatively
self-contained, we study the \kl{prefix relation} and prove in
\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
languages by the \kl{prefix relation}. In
\cref{infixes-regular:sec}, we
obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
specifically for \kl{bounded languages}
(\cref{bounded-language:thm}). Finally, 
we generalize this result to all
\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
in
(\cref{infix-amalgamation:thm}),
and provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered} by the \kl{infix relation} in
this context (\cref{infix-wqo-is-emptiness:thm}).

### 2024-10-18 src/prelims.tex
adding more and more todos

\paragraph*{Finite words.} In this paper, we use letters $\Sigma, \Gamma$ to
denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
some intuition on the decision problems, we will sometimes use the notion of
\intro{finite automata}, \intro{regular languages}, and \intro{Monadic Second
Order} ($\intro*\MSO$) logic on finite words, and assume is familiar with them.
We refer to the textbook of \cite{THOM97} for a detailed introduction to this
topic.

@@

\paragraph*{Ordinal Invariants.} An ordinal is a well-founded ordered set. We
use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to denote the
first infinite ordinal, i.e., the set of natural numbers with the usual
ordering. We assume superficial familiarity with ordinal arithmetic, and refer
to \cite{KUNEN80} for a detailed introduction to this domain.
Given a tree $T$ whose branches are all finite we can define
an ordinal $\alpha_T$ inductively as follows: if $T$ is a leaf then $\alpha_T =
0$, if $T$ has children $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i}
+ 1}{i \in \Nat}$. We say that $\alpha_T$ is the \emph{rank} of $T$.

>>>>

\paragraph*{Finite words.} In this paper, we use letters $\Sigma, \Gamma$ to
denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
some intuition on the decision problems, we will sometimes use the notion of
\intro{finite automata}, \intro{regular languages}, and Monadic Second Order
logic ($\intro*\MSO$) over finite words, and assume is familiar with them. We
refer to the textbook of \cite{THOM97} for a detailed introduction to this
topic.

@@

\paragraph*{Ordinal Invariants.} 
\label{ordinal-invariants:subsec} An ordinal is a well-founded ordered set. We
use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to denote the
first infinite ordinal, i.e., the set of natural numbers with the usual
ordering. We assume superficial familiarity with ordinal arithmetic, and refer
to \cite{KUNEN80} for a detailed introduction to this domain.
Given a tree $T$ whose branches are all finite we can define
an ordinal $\alpha_T$ inductively as follows: if $T$ is a leaf then $\alpha_T =
0$, if $T$ has children $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i}
+ 1}{i \in \Nat}$. We say that $\alpha_T$ is the \emph{rank} of $T$.

### 2024-10-18 src/trash.tex
adding more and more todos

    Now, thanks to \cref{inf-period-union-chains:lem},
    this happens to be a finite union of \kl{chains}
    for the \kl{infix relation}.
\end{proofof}

>>>>

    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
    relation (using \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
    for the \kl{suffix relation},
    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
    (using again \cref{inf-period-chain:lem}).
    As a consequence,
    $\InfPeriodChain{x}u \cdot \InfPeriodChain{y}$ is a finite union 
    of products of \kl{chains}, as desired.
\end{proofof}
